---
title: "qa_qc"
author: "Camila Vargas"
date: "1/19/2022"
output: html_document
---

Looking into Ana Sofia Guerra's data to make sure all data is clean, accurate and in the right format.

```{r setup, include=FALSE}

library(tidyverse)
library(freeR)
library(janitor)
library(lubridate)
library(here)

```


## Read data
```{r}

ac_b_morea <- read_csv(here::here("6.raw_data/acatri_behavior_moo2018.csv"))
ac_b_pal <- read_csv(here::here("6.raw_data/acatri_behavior_pal2017.csv"))

track_moo <- read_csv(here::here("6.raw_data/acatri_tracks_moo2018.csv"))
track_pal <- read_csv(here::here("6.raw_data/acatri_tracks_pal2017.csv"))


school_moo <- read_csv(here::here("6.raw_data/schooling_surveys_moo2018.csv"))

school_pal <- read_csv(here::here("6.raw_data/schooling_surveys_pal2017.csv"))

```



## Compare col names
```{r}
colnames(ac_b_morea)
colnames(ac_b_pal)
```

The behavior data between Moorea and Palmyra do not exactly match on col names

```{r}
colnames(track_moo)
olnames(track_pal)
```
The only difference is that Moorea has a Date column and Palmyra does not. Both have timestamps

```{r}
colnames(school_moo)
colnames(school_pal)

```

All the same. In both dataframes columns from 11 to 19 are NA columns.


## Are there NAs?
```{r}
freeR::complete(ac_b_morea) ##complete
freeR::complete(ac_b_pal) ##complete

freeR::complete(track_moo) ##complete
freeR::complete(track_pal) ##complete

freeR::complete(school_moo) ## all columns have 367 NAs except for 'schooling' that has 412
freeR::complete(school_pal) ## all columns have 2 NAs except for 'schooling' that has 170
```


## Check NAs are intentional NAs

```{r}
check_na_moo <- school_moo %>%
  filter_at(1:10, is.na) ## acrooss all variables. 367 NAs

##Opposite is to drop NAs/ We drop them from the first 8 columns because the column schooling does not have

no_na_moo <- school_moo %>% 
  drop_na(1:8)

check_na_pal <- school_pal %>%
  filter_at(1:10, is.na)

no_na_pal <- school_pal %>% 
  drop_na(1:8)
```

We need to remove observations with NAs across all columns


## Date and time format
```{r}
head(ac_b_morea)
class(ac_b_morea$Date) ## date is class character format: MM/DD/YY
class(ac_b_morea$time.stamp)
unique(ac_b_morea$time.stamp)

head(ac_b_pal)
class(ac_b_pal$Date) ## date is class character format: MM/DD/YY
unique(ac_b_pal$time.stamp) ## time.stamp are numbers from 1-60


head(track_moo)
class(track_moo$Date) ## date is class character format: MM/DD/YY
unique(track_moo$time.stamp) ## time.stamp are numbers from 1-60

head(track_pal) ## !!! NO Date!!!
unique(track_pal$time.stamp) ## time.stamp are numbers from 1-60


head(school_moo)
class(school_moo$Date) ## Date class character, format: DD-mmm-YY


head(school_pal)
class(school_pal$Date) ## date is class character format: MM/DD/YY

```

Need to format dates into YYYY-MM-DD

## Clean data
```{r}
## Behavior
clean_b_moo <- ac_b_morea %>% 
  mutate(Date = as.Date(Date, format = "%m/%d/%y"))

clean_b_pal <- ac_b_pal %>% 
  mutate(Date = as.Date(Date, format = "%m/%d/%y"))

##Tacking

clean_track_moo <- track_moo %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%y"))

##track_pal does not have date

clean_track_pal <- track_pal %>% 
  mutate(Date = str_remove(FollowID, "(?<=\\-).+"),
         Date = str_remove(Date, "\\-") %>% trimws(),
         Date = as.Date(Date, format = "%d%m%Y"))%>% 
  select(Island, FollowID, Date, everything())


## School
clean_school_moo <- school_moo %>% 
  select(1:10) %>% 
  drop_na(1:8) %>%
  mutate(Date = case_when(Date == "23-Aug-18" ~ "8/23/18",
                          T ~ Date),
         Date = as.Date(Date, format = "%m/%d/%y")) 
  
  
clean_school_pal <- school_pal %>%
  select(1:10) %>% 
  drop_na(1:8) %>%
  mutate(Date = as.Date(Date, format = "%m/%d/%y"))
  

```



## Check every category has the same format within a column
only character/categorical variables
```{r}
## Behavior

str(clean_b_moo)
unique(clean_b_moo$Island) #ok
unique(clean_b_moo$Observer) #ok
unique(clean_b_moo$behavior) #ok

str(clean_b_pal)
unique(clean_b_pal$Island) #ok
unique(clean_b_pal$Observer) ## Make sure each initial is from somone different. For Example is KD and KDM different?
unique(clean_b_pal$behavior) #ok

## Tacking

str(clean_track_moo)
unique(track_moo$Island) #ok
unique(clean_track_moo$Site) #ok

clean_track_pal <- track_pal

str(clean_track_pal)
unique(clean_track_pal$Island)
unique(clean_track_pal$Site) ## !!need definition!!

##School
str(clean_school_moo)
unique(clean_school_moo$Island) #ok
unique(clean_school_moo$Site) #ok
unique(clean_school_moo$Observer) #ok
unique(clean_school_moo$group) ## need definitions, make sure this are right
unique(clean_school_moo$species) ## need definitions, make sure this are right
unique(clean_school_moo$s.encounter.num) ## need definitions.


str(clean_school_pal)
unique(clean_school_pal$Island) #ok
unique(clean_school_pal$Site) #Need definitions
unique(clean_school_pal$Observer) #ok
unique(clean_school_pal$group) ## need definitions
unique(clean_school_pal$species) ## need definitions, make sure this are right
unique(clean_school_pal$s.encounter.num) ## need definitions.

```


## Save clean data

```{r}

## beahavior
write_csv(clean_b_moo, here::here("7.clean_data/acatri_behavior_moorea_2018.csv"))

write_csv(clean_b_pal, here::here("7.clean_data/acatri_behavior_palmyra_2017.csv"))

##Tracking
write_csv(clean_track_moo, here::here("7.clean_data/acatri_tracks_moorea_2018.csv"))

write_csv(clean_track_pal, here::here("7.clean_data/acatri_tracks_palmyra_2017.csv"))

## Schooling

write_csv(clean_school_moo, here::here("7.clean_data/schooling_surveys_moorea_2018.csv"))

write_csv(clean_school_pal, here::here("7.clean_data/schooling_surveys_palmyra_2017.csv"))



```


